# Lake Water Quality Profiles

---
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)

# clear environment
rm(list=ls())

# load packages
library(tidyverse)
library(googlesheets4)
library(lubridate)
library(readr)
library(readxl)
library(writexl)
library(hms)
library(plotly)
library(DT)
library(xlsx)
library(leaflet)
library(DT)
library(ggpubr)
library(plotrix)
library(packrat)

# set plotting themes

## geom_col plots theme
col_theme <- theme(axis.title = element_text(size = 14, face = "bold"),
                   strip.text = element_text(size = 14, face = "bold"),
                   legend.title = element_text(size = 14, face = "bold"),
                   legend.text = element_text(size = 14),
                   axis.text = element_text(size = 14))

## geom_points plots theme
points_theme <- theme(axis.title = element_text(size = 14, face = "bold"),
                   strip.text = element_text(size = 14, face = "bold"),
                   legend.title = element_text(size = 14, face = "bold"),
                   legend.text = element_text(size = 14),
                   axis.text = element_text(size = 9),
                   title = element_text(size = 18))

# function to exclude multiple items per column
'%ni%' <- Negate('%in%')

# clarify select function
select <- dplyr::select
```

Lake water quality profiles were collected at 1-2 month intervals at five sites at 1 meter depth intervals. Raw water quality field data is stored in a Google Sheet that can be viewed at <https://tinyurl.com/kwf-vogel-wqx-data>.

Data summaries and visualizations are provided for each site visit here.

```{r, include = FALSE}
# Import data from google sheet

# Note to self: once project is complete and no more data updates are needed, migrate data source from Google Sheet to csvs in the local project "data" folder

# recognize sheet title
lake_wqx_dat <- read_sheet("https://docs.google.com/spreadsheets/d/1lS9eJ7kX91IlYwSgiEQm-BQqif_7o5E4MzcBmqpwSoI/edit#gid=0", 
                           sheet = "Profile_Measurements") %>%
  # rename columns
  rename(site = `Site Name`,
         sample_date = `Sample Date`,
         sample_depth_m = `Depth (m)`,
         sample_time = `Sample Time`,
         DO_mgl = `DO (mg/L)`,
         turbidity_ntu = `Turbidity (NTU)`,
         temp_c = `Temp (C)`,
         spcond_uscm = `SpCond (uS/cm)`) %>%
  select(-Notes,-QC1,-`Data Entry`,-`QC2`) %>%
  
  # transform column formats
  transform(sample_time = hms::as_hms(sample_time)) %>%
  
  # convert data fromat from wide to long
  pivot_longer(cols = c("pH","DO_mgl","turbidity_ntu","temp_c","spcond_uscm"),
               names_to = "parameter",
               values_to = "val") %>%
  
  # remove missing values
  filter(!is.na(val))

```

```{r include = F}
# prep dataframe for plotting

# create parameter name replacements
param_names <- c("pH" = "pH",
                 "DO_mgl" = "DO\n(mg/L)",
                 "turbidity_ntu" = "Turbidity\n(NTU)",
                 "temp_c" = "Temperature\n(C)",
                 "spcond_uscm" = "Specific\nConductivity\n(uS/cm)")

# create site name replacements
site_names <- c("Vogel_Lake_Outlet" = "Vogel Lake Outlet", 
                "Vogel_Lake" = "Vogel Lake",        
                "North_Vogel_Lake" = "North Vogel Lake",
                "Bird_Pond" = "Bird Pond",
                "Vogel_Pond" = "Vogel Pond")

```

## January 2021

```{r ,include = FALSE}

# filter to january 2021 data data
jan21_lake_wqx_dat <- lake_wqx_dat %>%
  mutate(month = month(sample_date),
         year = year(sample_date)) %>%
  filter(month == "1",
         year == "2021")

# create plot
p <- jan21_lake_wqx_dat %>%
  ggplot() +
    geom_point(aes(sample_depth_m,val), size = 2) +
    geom_line(aes(sample_depth_m,val)) +

    coord_flip() +
    facet_grid(site ~ parameter, scales = "free_x", shrink = T, 
               labeller = labeller(parameter = param_names,
                                   site = site_names)) +
    ylab("") +
    xlab("Depth (m)") +
    theme_bw() +
    points_theme +
    scale_x_reverse() +
    scale_y_continuous(n.breaks = 5) +
    theme(panel.spacing = unit(1.2, "lines"),
          strip.text.y = element_text(angle = 360)) +
    ggtitle("January 2021 Water Quality All Sites")
```

```{r fig.width=10, fig.height=11}
p
```



## March 2021

```{r ,include = FALSE}

# filter to march 2021 data data
mar21_lake_wqx_dat <- lake_wqx_dat %>%
  mutate(month = month(sample_date),
         year = year(sample_date)) %>%
  filter(month == "3",
         year == "2021")

# create plot
p <- mar21_lake_wqx_dat %>%
  ggplot() +
    geom_point(aes(sample_depth_m,val), size = 2) +
    geom_line(aes(sample_depth_m,val)) +

    coord_flip() +
    facet_grid(site ~ parameter, scales = "free_x", shrink = T, 
               labeller = labeller(parameter = param_names,
                                   site = site_names)) +
    ylab("") +
    xlab("Depth (m)") +
    theme_bw() +
    points_theme +
    scale_x_reverse() +
    scale_y_continuous(n.breaks = 5) +
    theme(panel.spacing = unit(1.2, "lines"),
          strip.text.y = element_text(angle = 360)) +
    ggtitle("March 2021 Water Quality All Sites")
```

```{r fig.width=10, fig.height=11}
p
```

***

## Site Summaries

TBD
