# Lake Water Quality Profiles

---
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)

# clear environment
rm(list=ls())

# load packages
library(bookdown)
library(tidyverse)
library(googlesheets4)
library(lubridate)
library(readr)
library(readxl)
library(writexl)
library(hms)
library(plotly)
library(DT)
library(xlsx)
library(leaflet)
library(DT)
library(ggpubr)
library(plotrix)
library(packrat)
library(magrittr)

# set plotting themes

## geom_col plots theme
col_theme <- theme(axis.title = element_text(size = 14, face = "bold"),
                   strip.text = element_text(size = 14, face = "bold"),
                   legend.title = element_text(size = 14, face = "bold"),
                   legend.text = element_text(size = 14),
                   axis.text = element_text(size = 14))

## geom_points plots theme
points_theme <- theme(axis.title = element_text(size = 14, face = "bold"),
                   strip.text = element_text(size = 14, face = "bold"),
                   legend.title = element_text(size = 14, face = "bold"),
                   legend.text = element_text(size = 14),
                   axis.text = element_text(size = 11, face = "bold"),
                   title = element_text(size = 18))

# function to exclude multiple items per column
'%ni%' <- Negate('%in%')

# clarify select function
select <- dplyr::select
```

Lake water quality profiles were collected at 1-2 month intervals at five sites at 1 meter depth intervals. Hydrolab MS5 sondes were descended from an anchored boat centered over the deepest point of each lake. 

Raw water quality field data is stored in a Google Sheet that can be viewed at <https://tinyurl.com/kwf-vogel-wqx-data>.

Data visualizations are provided for each site visit here.

```{r, include = FALSE}
# Import data from google sheet

# Note to self: once project is complete and no more data updates are needed, migrate data source from Google Sheet to csvs in the local project "data" folder

# recognize sheet title and download
lake_wqx_dat <- read_sheet("https://docs.google.com/spreadsheets/d/1lS9eJ7kX91IlYwSgiEQm-BQqif_7o5E4MzcBmqpwSoI/edit#gid=0", 
                           sheet = "Lake_Profile_Measurements") %>%
  # rename columns
  rename(site = `Site Name`,
         sample_date = `Sample Date`,
         sample_depth_m = `Depth (m)`,
         sample_time = `Sample Time`,
         DO_mgl = `DO (mg/L)`,
         turbidity_ntu = `Turbidity (NTU)`,
         temp_c = `Temp (C)`,
         spcond_uscm = `SpCond (uS/cm)`) %>%
  select(-Notes,-QC1,-`Data Entry`,-`QC2`) %>%
  
  # transform column formats
  transform(sample_time = hms::as_hms(sample_time),
            pH = as.character(pH),
            DO_mgl = as.character(DO_mgl),
            turbidity_ntu = as.character(turbidity_ntu),
            temp_c = as.character(temp_c),
            spcond_uscm = as.character(temp_c)) %>%
  
  # convert data format from wide to long
  pivot_longer(cols = c("pH","DO_mgl","turbidity_ntu","temp_c","spcond_uscm"),
               names_to = "parameter",
               values_to = "val") %>%
  
  # remove missing values
  filter(!is.na(val)) %>%
  
  # remove values flagged with an asterisk
  # note that the asterisk is a special character, requires // escape for spring match
  filter(!str_detect(val, "\\*")) %>%
  
  # convert values back to numeric type after filtering out flagged data
  transform(val = as.numeric(val)) %>%
  
  # add new columns
  mutate(month = month(sample_date))

```

```{r include = F}
# prep dataframe for plotting

# create parameter name replacements
param_names <- c("pH" = "pH",
                 "DO_mgl" = "DO\n(mg/L)",
                 "turbidity_ntu" = "Turbidity\n(NTU)",
                 "temp_c" = "Temperature\n(C)",
                 "spcond_uscm" = "Specific\nConductivity\n(uS/cm)")

# create site name replacements
site_names <- c("Vogel_Lake_Outlet" = "Vogel\nLake\nOutlet", 
                "Vogel_Lake" = "Vogel\nLake",        
                "North_Vogel_Lake" = "North\nVogel\nLake",
                "Bird_Pond" = "Bird\nPond",
                "Vogel_Pond" = "Vogel\nPond")

```



## Data by Month

### January 2021

Jan 22, 2021

```{r jan21-wqx, echo = FALSE, fig.cap="January 2021", fig.width=10, fig.height=11}

# filter to january 2021 data data
jan21_lake_wqx_dat <- lake_wqx_dat %>%
  mutate(month = month(sample_date),
         year = year(sample_date)) %>%
  filter(month == "1",
         year == "2021")

# create plot
jan21_lake_wqx_dat %>%
  ggplot() +
    geom_point(aes(sample_depth_m,val), size = 2) +
    geom_line(aes(sample_depth_m,val)) +

    coord_flip() +
    facet_grid(site ~ parameter, scales = "free_x", shrink = T, 
               labeller = labeller(parameter = param_names,
                                   site = site_names)) +
    ylab("") +
    xlab("Depth (m)") +
    theme_bw() +
    points_theme +
    scale_x_reverse() +
    scale_y_continuous(n.breaks = 5) +
    theme(panel.spacing = unit(1.2, "lines"),
          strip.text.y = element_text(angle = 360)) +
    ggtitle("January 2021 Water Quality All Sites")
```




### March 2021

March 23, 2021

```{r , echo = FALSE, fig.cap = "March 2021", fig.width=10, fig.height=11}

# filter to march 2021 data data
mar21_lake_wqx_dat <- lake_wqx_dat %>%
  mutate(month = month(sample_date),
         year = year(sample_date)) %>%
  filter(month == "3",
         year == "2021")

# create plot
mar21_lake_wqx_dat %>%
  ggplot() +
    geom_point(aes(sample_depth_m,val), size = 2) +
    geom_line(aes(sample_depth_m,val)) +

    coord_flip() +
    facet_grid(site ~ parameter, scales = "free_x", shrink = T, 
               labeller = labeller(parameter = param_names,
                                   site = site_names)) +
    ylab("") +
    xlab("Depth (m)") +
    theme_bw() +
    points_theme +
    scale_x_reverse() +
    scale_y_continuous(n.breaks = 5) +
    theme(panel.spacing = unit(1.2, "lines"),
          strip.text.y = element_text(angle = 360)) +
    ggtitle("March 2021 Water Quality All Sites")
```


### May 2021
May 25, 2021

```{r ,echo = FALSE, fig.cap = "May 2021", fig.width=10, fig.height=11}

# filter to may 2021 data data
may21_lake_wqx_dat <- lake_wqx_dat %>%
  mutate(month = month(sample_date),
         year = year(sample_date)) %>%
  filter(month == "5",
         year == "2021")

# create plot
may21_lake_wqx_dat %>%
  ggplot() +
    geom_point(aes(sample_depth_m,val), size = 2) +
    geom_line(aes(sample_depth_m,val)) +
    coord_flip() +
    facet_grid(site ~ parameter, scales = "free_x", shrink = T, 
               labeller = labeller(parameter = param_names,
                                   site = site_names)) +
    ylab("") +
    xlab("Depth (m)") +
    theme_bw() +
    points_theme +
    scale_x_reverse() +
    scale_y_continuous(n.breaks = 5) +
    theme(panel.spacing = unit(1.2, "lines"),
          strip.text.y = element_text(angle = 360)) +
    ggtitle("May 2021 Water Quality All Sites")
```



### June 2021
June 29, 2021

```{r ,echo = FALSE, fig.cap = "June 2021", fig.width=10, fig.height=11}

# filter to june 2021 data data
jun21_lake_wqx_dat <- lake_wqx_dat %>%
  mutate(month = month(sample_date),
         year = year(sample_date)) %>%
  filter(month == "6",
         year == "2021")

# create plot
jun21_lake_wqx_dat %>%
  ggplot() +
    geom_point(aes(sample_depth_m,val), size = 2) +
    geom_line(aes(sample_depth_m,val)) +

    coord_flip() +
    facet_grid(site ~ parameter, scales = "free_x", shrink = T, 
               labeller = labeller(parameter = param_names,
                                   site = site_names)) +
    ylab("") +
    xlab("Depth (m)") +
    theme_bw() +
    points_theme +
    scale_x_reverse() +
    scale_y_continuous(n.breaks = 5) +
    theme(panel.spacing = unit(1.2, "lines"),
          strip.text.y = element_text(angle = 360)) +
    ggtitle("June 2021 Water Quality All Sites")
```

Note: turbidity data is absent from June 2021 sampling due to an equipment issue.

<br>

### August 2021
August 3, 2021

```{r ,echo = FALSE, fig.cap = "August 2021", fig.width=10, fig.height=11}

# filter to aug 2021 data data
aug21_lake_wqx_dat <- lake_wqx_dat %>%
  mutate(month = month(sample_date),
         year = year(sample_date)) %>%
  filter(month == "8",
         year == "2021")

# create plot
aug21_lake_wqx_dat %>%
  ggplot() +
    geom_point(aes(sample_depth_m,val), size = 2) +
    geom_line(aes(sample_depth_m,val)) +

    coord_flip() +
    facet_grid(site ~ parameter, scales = "free_x", shrink = T, 
               labeller = labeller(parameter = param_names,
                                   site = site_names)) +
    ylab("") +
    xlab("Depth (m)") +
    theme_bw() +
    points_theme +
    scale_x_reverse() +
    scale_y_continuous(n.breaks = 5) +
    theme(panel.spacing = unit(1.2, "lines"),
          strip.text.y = element_text(angle = 360)) +
    ggtitle("August 2021 Water Quality All Sites")
```

<br>

### September 2021
September 15, 2021

```{r ,echo = FALSE, fig.cap = "September 2021", fig.width=10, fig.height=11}

# filter to sept 2021 data data
sep21_lake_wqx_dat <- lake_wqx_dat %>%
  mutate(month = month(sample_date),
         year = year(sample_date)) %>%
  filter(month == "9",
         year == "2021")

# create plot
sep21_lake_wqx_dat %>%
  ggplot() +
    geom_point(aes(sample_depth_m,val), size = 2) +
    geom_line(aes(sample_depth_m,val)) +

    coord_flip() +
    facet_grid(site ~ parameter, scales = "free_x", shrink = T, 
               labeller = labeller(parameter = param_names,
                                   site = site_names)) +
    ylab("") +
    xlab("Depth (m)") +
    theme_bw() +
    points_theme +
    scale_x_reverse() +
    scale_y_continuous(n.breaks = 5) +
    theme(panel.spacing = unit(1.2, "lines"),
          strip.text.y = element_text(angle = 360)) +
    ggtitle("September 2021 Water Quality All Sites")
```

<br>

### December 2021
December 2 and 3, 2021

```{r ,echo = FALSE, fig.cap = "December 2021", fig.width=10, fig.height=11}

# filter to dec 2021 data 
dec21_lake_wqx_dat <- lake_wqx_dat %>%
  mutate(month = month(sample_date),
         year = year(sample_date)) %>%
  filter(month == "12",
         year == "2021")

# create plot
dec21_lake_wqx_dat %>%
  ggplot() +
    geom_point(aes(sample_depth_m,val), size = 2) +
    geom_line(aes(sample_depth_m,val)) +

    coord_flip() +
    facet_grid(site ~ parameter, scales = "free_x", shrink = T, 
               labeller = labeller(parameter = param_names,
                                   site = site_names)) +
    ylab("") +
    xlab("Depth (m)") +
    theme_bw() +
    points_theme +
    scale_x_reverse() +
    scale_y_continuous(n.breaks = 5) +
    theme(panel.spacing = unit(1.2, "lines"),
          strip.text.y = element_text(angle = 360)) +
    ggtitle("December 2021 Water Quality All Sites")
```

<br>


### December 2021
December 2 and 3, 2021

```{r ,echo = FALSE, fig.cap = "January 2022", fig.width=10, fig.height=11}

# filter to jan 2022 data 
jan22_lake_wqx_dat <- lake_wqx_dat %>%
  mutate(month = month(sample_date),
         year = year(sample_date)) %>%
  filter(month == "1",
         year == "2022")

# create plot
jan22_lake_wqx_dat %>%
  ggplot() +
    geom_point(aes(sample_depth_m,val), size = 2) +
    geom_line(aes(sample_depth_m,val)) +

    coord_flip() +
    facet_grid(site ~ parameter, scales = "free_x", shrink = T, 
               labeller = labeller(parameter = param_names,
                                   site = site_names)) +
    ylab("") +
    xlab("Depth (m)") +
    theme_bw() +
    points_theme +
    scale_x_reverse() +
    scale_y_continuous(n.breaks = 5) +
    theme(panel.spacing = unit(1.2, "lines"),
          strip.text.y = element_text(angle = 360)) +
    ggtitle("January 2022 Water Quality All Sites")
```



***

## Overall Data

#### Dissolved Oxygen

```{r , echo = FALSE, fig.cap = "Dissolved Oxygen (mg/L)", fig.width=10, fig.height=8}
# plot
lake_wqx_dat %>%
  filter(parameter == "DO_mgl") %>%
  ggplot(aes(as.factor(sample_date),sample_depth_m)) +
  geom_tile(aes(fill = val)) +
  scale_y_reverse() +
  facet_grid(. ~ site, labeller = labeller(site = site_names)) +
    ylab("Depth (m)") +
    xlab("") +
    theme_bw() +
   points_theme +
   theme(axis.text.x=element_text(angle=90, vjust=.5)) +
   scale_fill_viridis_c(name="mg/L") +
  ggtitle ("Dissolved Oxygen")

```



<br>

#### pH

```{r , echo = FALSE, fig.cap = "pH", fig.width=10, fig.height=8}
# plot
lake_wqx_dat %>%
  filter(parameter == "pH") %>%
  ggplot(aes(as.factor(sample_date),sample_depth_m)) +
  geom_tile(aes(fill = val)) +
  scale_y_reverse() +
  facet_grid(. ~ site, labeller = labeller(site = site_names)) +
    ylab("Depth (m)") +
    xlab("") +
    theme_bw() +
   points_theme +
   theme(axis.text.x=element_text(angle=90, vjust=.5)) +
   scale_fill_viridis_c(name="value") +
  ggtitle ("pH")

```

Note: pH values are excluded from the May and June 2021 site visits due to an instrument error.

<br>

#### Turbidity

```{r , echo = FALSE, fig.cap = "Turbidity (NTU)", fig.width=10, fig.height=8}

# plot
lake_wqx_dat %>%
  filter(parameter == "turbidity_ntu") %>%
  ggplot(aes(as.factor(sample_date),sample_depth_m)) +
  geom_tile(aes(fill = val)) +
  scale_y_reverse() +
  facet_grid(. ~ site, labeller = labeller(site = site_names)) +
    ylab("Depth (m)") +
    xlab("") +
    theme_bw() +
   points_theme +
   theme(axis.text.x=element_text(angle=90, vjust=.5)) +
   scale_fill_viridis_c(name="NTU") +
  ggtitle ("Turbidity")

```

Note: some turbidity values near the benthic surface of each site visit are not displayed in the above plot in order to improve visualization.

<br>

#### Conductivity

```{r , echo = FALSE, fig.cap = "Conductivity (uS/cm)", fig.width=10, fig.height=8}
# plot
lake_wqx_dat %>%
  filter(parameter == "spcond_uscm") %>%
  ggplot(aes(as.factor(sample_date),sample_depth_m)) +
  geom_tile(aes(fill = val)) +
  scale_y_reverse() +
  facet_grid(. ~ site, labeller = labeller(site = site_names)) +
    ylab("Depth (m)") +
    xlab("") +
    theme_bw() +
   points_theme +
   theme(axis.text.x=element_text(angle=90, vjust=.5)) +
   scale_fill_viridis_c(name="uS/cm") +
  ggtitle ("Conductivity")

```


<br>

#### Water Temperature

```{r , echo = FALSE, fig.cap = "Water Temperature (C)", fig.width=10, fig.height=8}
# plot
lake_wqx_dat %>%
  filter(parameter == "temp_c") %>%
  ggplot(aes(as.factor(sample_date),sample_depth_m)) +
  geom_tile(aes(fill = val)) +
  scale_y_reverse() +
  facet_grid(. ~ site, labeller = labeller(site = site_names)) +
    ylab("Depth (m)") +
    xlab("") +
    theme_bw() +
   points_theme +
   theme(axis.text.x=element_text(angle=90, vjust=.5)) +
   scale_fill_viridis_c(name="°C") +
  ggtitle ("Temperature")

```



***

## Site Summaries

A table of summary statistics for water quality parameters at each site will be provided here.
